---
title: "Mapcode"
output: html_document
date: "2026-03-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sandprawn distribution and protection level

Creating an interactive map of *Kraussillichirus kraussi* observations in South Africa colour coded by ecosystem protection level.

#### Reading in coastal protection data.

```{r}
library(sf)
cep <- st_read("NBA_benthic_and_coastal_protection.shp")
st_crs(cep)
head(cep)
plot(cep["PA_level"])
```

#### Reading in the data from iNaturalist.

```{r}
library(rinat)
sp <- get_inat_obs(taxon_name = "Kraussillichirus kraussi", bounds = c(-35, 17.5, -26.8, 33), maxresults = 1000)
head(sp)
```

Filtering the results for research grade observations and positional accuracy. Converting the dataframe to a spatial object.

```{r}
library(tidyverse)
sp <- sp %>% filter(positional_accuracy<46 & latitude<0 & !is.na(latitude) & quality_grade == "research")

#converting into spatial object

sp <- st_as_sf(sp, coords = c("longitude", "latitude"), crs = 4326)

```

#### Intersecting points with polygons.

```{r}

sp  <- st_make_valid(sp)
cep <- st_make_valid(cep)

sp2 <- st_intersection(sp, cep)
```

Colour coding points in an interactive map with clickable url links.

```{r}
library(leaflet)
library(leafpop)
library(tidyverse)

#creating a colour palette
pal <- colorFactor(palette = c("darkblue", "green", "purple", "darkgreen", "orange"), domain = sp2$PA_level)
#clickable urls 
lsp2 <- sp2 %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

leaflet(sp2) %>% addTiles() %>% addCircleMarkers(radius = 5, fillColor = ~pal(PA_level), fillOpacity = 0.8, stroke = FALSE, popup = popupTable(lsp2, zcol = c("user_login", "click_url"))) %>% addLegend( "bottomright", pal = pal, values = ~PA_level, title = "Protection Level")
```
