---
title: "Mapcode"
output: html_document
date: "2026-03-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sandprawn distribution and protection level

Creating an interactive map of *Kraussillichirus kraussi* observations in South Africa colour coded by ecosystem protection level.

Reading in coastal protection data.

```{r}
library(sf)
cep <- st_read("NBA_benthic_and_coastal_protection.shp")
```

Reading in the data from iNaturalist.

```{r}
library(rinat)
sp <- get_inat_obs(taxon_name = "Kraussillichirus kraussi", bounds = c(-35, 17.5, -26.8, 33), maxresults = 1000)
head(sp)
```

Filtering the results for research grade observations and positional accuracy. Converting the dataframe to a spatial object.

```{r}
library(tidyverse)
sp <- sp %>% filter(positional_accuracy<46 & latitude<0 & !is.na(latitude) & quality_grade == "research")

#converting into spatial object

sp <- st_as_sf(sp, coords = c("longitude", "latitude"), crs = 4326)

```

Creating an interactive map.

```{r}

library(leaflet)
library(htmltools)

leaflet() %>% addTiles(group = "Default") %>% addCircleMarkers(data = sp, group = "Kraussillichirus kraussi", radius = 3, color = "green") 
```



